### 2013-07-22 doc/dissertation/document/chapters/mc_background/mc_background.tex
adding more to chapter summaries
fixing minres -> minimal residual

\subsection{Alternative Fixed Point Iterations}
\label{subsubsec:alternative_fixed_point}
In addition to the basic Richardson iteration, the MCSA algorithm
presented in Eq~(\ref{eq:mcsa}) can be used to accelerate any fixed
point iteration which only depends on the state (typically the
residual) of the last iteration. As outlined in
Appendix~\ref{ch:linear_problem}, subspace methods take on a general
form using the Petrov-Galerkin conditions as constraints for
extracting a correction from the search subspace as given by
Eq~(\ref{eq:linear_projection_iteration}). Those that are
one-dimensional may be used with fixed-point MCSA as they only depend
the previous iteration for information. As an example, for
positive-definite but not necessarily symmetric problems, the minimal
residual (MINRES) iteration \citep{saad_iterative_2003} can be used in
conjunction with MCSA where the residual vector, $\mathbf{r}$, defines
the search subspace and the action of the linear operator on the
residual vector, $\mathbf{A}\mathbf{r}$, defines the constraint
subspace. Using this, we can then define an MCSA scheme that
accelerates the MINRES iteration:
\begin{subequations}
  \begin{gather}
    \alpha = \frac{\langle \mathbf{A}\mathbf{r}^k, \mathbf{r}^k
      \rangle}{\langle \mathbf{A}\mathbf{r}^k, \mathbf{A}\mathbf{r}^k
      \rangle}\:, \\
    \ve{x}^{k+1/2} = \ve{x}^k + \alpha \ve{r}^k\:,
    \label{eq:minres_correction}\\
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\ 
    \ve{A}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:,\\ 
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta\ve{x}^{k+1/2}\:,\\ 
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:.
  \end{gather}
  \label{eq:mcsa_minres}
\end{subequations}
where $\alpha$ is an optimal extrapolation parameter generated from
the constraints. Interestingly, this scheme is nearly identical to the
Richardson iteration version in Eq~(\ref{eq:mcsa}) except now the
additional extrapolation parameter, $\alpha$, is applied to the
residual correction in Eq~(\ref{eq:minres_correction}) as a means of
selecting a more optimal search direction at each iteration,
potentially further accelerating convergence.

@@

\begin{itemize}
\item MCSA has been generalized for all discrete linear transport
  problems
\item MCSA can be preconditioned using a general left/right scheme but
  requires the explicit inversion of the preconditioners and
  construction of the composite linear operator
\item The adjoint Neumann-Ulam method performs better when used with
  MCSA compared to the forward method
\item MCSA has superior performance to Halton's residual Monte Carlo
  method
\item The Richardson iteration in MCSA behaves as a smoother for the
  stochastic Monte Carlo error
\item MCSA iterative performance when used with the expected value
  estimator is insensitive to histories per iteration
\item MCSA iterative performance is insensitive to the stochastic
  history weight cutoff
\item Variance reduction with artificial absorption can enhance time
  to solution
\item The reduced domain approximation was introduced to mitigate the
  consequences of composite linear operator generation
\end{itemize}

>>>>

\subsection{Alternative Fixed Point Iterations}
\label{subsubsec:alternative_fixed_point}
In addition to the basic Richardson iteration, the MCSA algorithm
presented in Eq~(\ref{eq:mcsa}) can be used to accelerate any fixed
point iteration which only depends on the state (typically the
residual) of the last iteration. As outlined in
Appendix~\ref{ch:linear_problem}, subspace methods take on a general
form using the Petrov-Galerkin conditions as constraints for
extracting a correction from the search subspace as given by
Eq~(\ref{eq:linear_projection_iteration}). Those that are
one-dimensional may be used with fixed-point MCSA as they only depend
the previous iteration for information. As an example, for
positive-definite but not necessarily symmetric problems, the minimal
residual iteration \citep{saad_iterative_2003} can be used in
conjunction with MCSA where the residual vector, $\mathbf{r}$, defines
the search subspace and the action of the linear operator on the
residual vector, $\mathbf{A}\mathbf{r}$, defines the constraint
subspace. Using this, we can then define an MCSA scheme that
accelerates the minimal residual iteration:
\begin{subequations}
  \begin{gather}
    \alpha = \frac{\langle \mathbf{A}\mathbf{r}^k, \mathbf{r}^k
      \rangle}{\langle \mathbf{A}\mathbf{r}^k, \mathbf{A}\mathbf{r}^k
      \rangle}\:, \\
    \ve{x}^{k+1/2} = \ve{x}^k + \alpha \ve{r}^k\:,
    \label{eq:min_res_correction}\\
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\ 
    \ve{A}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:,\\ 
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta\ve{x}^{k+1/2}\:,\\ 
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:.
  \end{gather}
  \label{eq:mcsa_min_res}
\end{subequations}
where $\alpha$ is an optimal extrapolation parameter generated from
the constraints. Interestingly, this scheme is nearly identical to the
Richardson iteration version in Eq~(\ref{eq:mcsa}) except now the
additional extrapolation parameter, $\alpha$, is applied to the
residual correction in Eq~(\ref{eq:min_res_correction}) as a means of
selecting a more optimal search direction at each iteration,
potentially further accelerating convergence.

@@

\begin{itemize}
\item Monte Carlo synthetic acceleration (MCSA) has been generalized
  for all discrete linear transport problems
\item MCSA can be preconditioned using a general left/right scheme but
  requires the explicit inversion of the preconditioners and
  construction of the composite linear operator
\item The adjoint Neumann-Ulam method performs better when used with
  MCSA compared to the forward method
\item MCSA has superior performance to Halton's residual Monte Carlo
  method
\item The Richardson iteration in MCSA behaves as a smoother for the
  stochastic Monte Carlo error
\item MCSA iterative performance when used with the expected value
  estimator is insensitive to histories per iteration
\item MCSA iterative performance is insensitive to the stochastic
  history weight cutoff
\item Variance reduction with artificial absorption can enhance time
  to solution
\item The reduced domain approximation was introduced to mitigate the
  consequences of composite linear operator generation
\end{itemize}
